Grove: A Convergent Collaborative Structure-Editor Calculus

### 2021-02-22 grove-paper.tex
edits with Cyrus

\newcommand{\Edge}[1]{Edge~#1}
\newcommand{\Vertex}[1]{Vertex~#1}
\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{\id{#1}}}}}
\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{\id{#1}}}}}
\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
\newcommand{\otherVertexVskip}{\vskip0.5em}

@@

A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
where $\E = \U \times \V \times \P \times \V$,
unique IDs are drawn from some suitable set $\U$ equipped with a total ordering,
vertices are drawn from $\V$,
positions are drawn from $\P$,
and edge states are drawn from $\Sigma$, all defined below.

@@

Let us now examine each of the four components in more detail.
The first component $e_r$ is the expression derived from the root vertex $v_0$.
\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
There is one rule for each constructor form.
For example, the expression derived from $v_{\id{u}} = (u, \ExpPlus)$ is $\expr'(v_{\id{u}}, \PlusLeft) \text{+}^{\id{u}} \expr'(v_{\id{u}}, \PlusRight)$.
In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
An empty hole $\hole$ appears if there are no children at the given position.
A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
If there is one child, the expression is determined recursively unless that child appears in $MP$ or $W$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.

@@

The fourth component $W$ is the set of expressions derived from the remaining vertices, which form cyclic structures we call wreaths, where each vertex has exactly one parent.
In $W$, $\ancestors(v)$ is the set of vertices for which there is a path to $v$.
Note that $v = \min(\ancestors(v))$, where $\min(\{v_{\id{u_1}}, \ldots, v_{\id{u_n}}\}) = v_{\id{u_k}}$ such that $u_k = \min(\{u_{\id{1}}, \ldots, u_{\id{n}}\})$, holds only if $v$ is part of a cycle.

@@

These first two components do not account for vertices with no parents. We
include the expressions corresponding to these vertices in the third component
$NP$.

We've accounted for single-parented vertices connected to the root,
single-parented vertices connected to multi-parented vertices, and vertices with
no parents. This leaves only single-parented vertices that are not connected to
the root or any multi-parented vertices. These vertices form cyclic structures
we call wreaths.

% A wreath is a cycle, potentially with acyclic branches. Wreaths are connected
% graphs where each vertex has exactly one incoming edge. The fourth component $W$
% is the set of expressions determined by extracting wreaths from any vertices not
% in the other three components.

% expression = tree

% (show what a wreath looks like)

>>>>

\newcommand{\Edge}[1]{Edge~#1}
\newcommand{\Vertex}[1]{Vertex~#1}
\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}
\newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\curlyveedownarrow_{#1}}}}
\newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\rcirclearrowleft_{#1}}}}
\newcommand{\orphanVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\ndownarrow_{#1}}}}}
\newcommand{\parens}[1]{\textcolor{gray}{(}#1\textcolor{gray}{)}}
\newcommand{\otherVertexVskip}{\vskip0.5em}

@@

A graph $G : \E \rightarrow \Sigma$ is a function from edges to edge states,
where $\E = \U \times \V \times \P \times \V$,
unique IDs are drawn from some suitable set $\U$ equipped with a total ordering $\leq$,
vertices are drawn from $\V$,
positions are drawn from $\P$,
and edge states are drawn from $\Sigma$, all defined below.

@@

Let us now examine each of the four components in more detail.
The first component $e_r$ is the expression derived from the root vertex $v_0$.
\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
There is one rule for each constructor form.
For example, the expression derived from ${+}_{\id{38}} = (38, \ExpPlus)$ is $\expr'({+}_{\id{38}}, \PlusLeft) \text{+}^{\id{38}} \expr'({+}_{\id{38}}, \PlusRight)$.
In each child position, $\expr'(v, p)$ determines the sub-expression derived from the children of $v$ at position $p$.
An empty hole $\hole$ appears if there are no children at the given position.
A conflict hole $\conflictHole{e_1,\ldots,e_n}$ appears if there are multiple children.
If there is one child, the expression is determined recursively unless that child appears in $MP$ or $W$, discussed below, in which case we leave a reference, $\multiVertex{u}$ or $\cycleVertex{u}$ respectively.
For the $\ExpLam$ case, we define $\patt(q)$ and $\type(\tau)$ similarly.

@@

The fourth component $W$ is the set of expressions derived from the remaining vertices, which form cyclic structures we call wreaths, where each vertex has exactly one parent.
Every wreath consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
We need to break the cycle to be able to turn a wreath into a tree.
To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id according to the total ordering on $\U$.
Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
every vertex on the cycle is its own ancestor.

@@
